---
title: "Data_Summary"
author: "Jessica French"
date:  "`r Sys.Date()`"
output: html_document
pandoc_args: ["+RTS",  "-M1024M",  "-RTS"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This notebook will show the locations that we have nutrient data for and a summary of other data we have aquired.

```{r label = "Load Packages"}
# Load in libraries
library(tidyverse)
library(sf)
library(tmap)
library(ncdf4)
library(terra)
library(janitor)
```

## Read in the boundary data

```{r label = "Read in AOI"}
# Read in the shape file of the area of interest
AOI <- st_read("/Users/jfrench/Documents/MEDS/Capstone/DATA/AOI_SBchannel_shp/AOI_SBchannel.shp")

# Read in expanded AOI for filtering large data sets
Expanded_AOI <- st_read("/Users/jfrench/Documents/MEDS/Capstone/DATA/expanded_AOI/Expaned_AOI_SBchannel.shp")
```

```{r label = "Read in Boundaries"}
# Read in state and federal boundaries
federal_boundaries <- st_read("/Users/jfrench/Documents/MEDS/Capstone/DATA/maritime_boundaries/federal_boundaries/USMaritimeLimitsNBoundaries.shp") |> 
  st_filter(Expanded_AOI)

# Read in California MPA boundaries
MPA_boundaries <- st_read("/Users/jfrench/Documents/MEDS/Capstone/DATA/MPAs/mpa_ca_4326 (1)/mpa_ca_4326Polygon.shp")|> 
  st_filter(Expanded_AOI)
```

## Read in the nutrient data

-   CalCOFI data has nitrogen species and phosphorous goes back past 2014

```{r label = "Read in CalCOFI"}
# Read in CalCOFI cast data
CalCOFI_cast <- read_csv("/Users/jfrench/Documents/MEDS/Capstone/DATA/Nutrients/CalCOFI_Database_194903-202001_csv_22Sep2021/194903-202001_Cast.csv") |> 
  filter(Year >= 2014) |> 
  select(c("Cst_Cnt", "Quarter", "Year", "Lat_Dec", "Lon_Dec")) 

# Read in the CalCOFI bottle data
CalCOFI_bottle <- read_csv("/Users/jfrench/Documents/MEDS/Capstone/DATA/Nutrients/CalCOFI_Database_194903-202001_csv_22Sep2021/194903-202001_Bottle.csv", locale=locale(encoding="latin1")) |> 
  right_join(CalCOFI_cast, by = "Cst_Cnt")|> 
  st_as_sf(coords = c(lon = "Lon_Dec", lat = "Lat_Dec"), crs = "EPSG: 4326") |> 
  st_filter(Expanded_AOI)
```

-   The LTER Biomass and environmental drivers dataset only has nitrogen and wave action
-   Goes back past 2014

```{r label = "Read in LTER Nitrogen"}
# Read in the LTER Biomass and Nitrogen Data Set
LTER_nutrients_biomass <- read_csv("/Users/jfrench/Documents/MEDS/Capstone/DATA/Nutrients/Nitrate_LTER/kelp_no3_waves_quarterly_long (1).csv")|> 
  filter(year >= 2014)

# Read in the LTER sampling locations
LTER_locations <- read_csv("/Users/jfrench/Documents/MEDS/Capstone/DATA/Nutrients/Nitrate_LTER/locations (1).csv")

# Combine the locations to the nutrient data
lter_data_geom <- left_join(LTER_nutrients_biomass, LTER_locations, by = "site_id") |> 
  st_as_sf(coords = c(lon = "lon", lat = "lat"), crs = "EPSG:4326")
```

-   Plumes and Blooms Data has nitrogen species and phosphorous
-   From before 1995 to 2016

```{r label = "Read in Plumes and Blooms"}
# Read in the plumes and blooms data
plumes_blooms <- read_csv("/Users/jfrench/Documents/MEDS/Capstone/DATA/Nutrients/plumes_blooms/Plumes_and_Blooms_bottle_file_data.csv", na = c('-999.00000', '-999', ""))|>
  mutate("DATE" = lubridate::ymd(DATE),
         "LON" = LON*-1)|>
  filter(DATE >= 2014-01-01,
         LON < 100,
         LAT >= 30 & LAT <= 40) |>
  st_as_sf(coords = c(lon = "LON", lat = "LAT"), crs = "EPSG:4326")
```

-   USGS Bottle Data from OCT and NOV 2018 and 2019

```{r label = "Read in USGS Bottle"}
# Read in USGS bottle data 2018
USGS_2018 <- read_csv("/Users/jfrench/Documents/MEDS/Capstone/DATA/usgs_bottle_fall/SH-18-12_BTL_CTD_data_v3.0.csv") |> 
  st_as_sf(coords = c("Longitude", "Latitude"),
           crs = 4326) |> 
  st_filter(Expanded_AOI)

# Read in USGS bottle data 2019
USGS_2019 <- read_csv("/Users/jfrench/Documents/MEDS/Capstone/DATA/usgs_bottle_fall/RL-19-05_BTL_CTD_data v2.0.csv") |> 
  st_as_sf(coords = c("Longitude", "Latitude"),
           crs = 4326) |> 
  st_filter(Expanded_AOI)

```

-   LTER Water Chemistry data

```{r label = "Read in LTER Water Chemistry"}
# Read in the Bottle file from LTER

LTER_bottle <- read_csv2("/Users/jfrench/Documents/MEDS/Capstone/DATA/Nutrients/LTER_bottle/LTER_monthly_bottledata_20220930.txt") |> 
  clean_names() |> 
  filter(longitude_e < 0) |> 
  select(po4_umol_l, no2_no3_umol_l, latitude_n, longitude_e)

stringi::stri_sub(LTER_bottle$latitude_n, 3, 2) = '.' 
stringi::stri_sub(LTER_bottle$longitude_e, 5, 4) <- '.'
#Encoding(LTER_bottle$comments <- "UTF-8")

LTER_bottle <- LTER_bottle |> 
  mutate("latitude_n" = as.numeric(latitude_n),
         "longitude_e" = as.numeric(longitude_e)) |> 
  st_as_sf(coords = c("longitude_e", "latitude_n"), 
           crs = 4326) 
```

```{r label = "Map"}
# Test map
tmap_mode(mode = "view")

tm_shape(AOI) +
  tm_borders(col = "red") +
  tm_shape(federal_boundaries) +
  tm_lines() +
  tm_shape(MPA_boundaries) +
  tm_borders(col = "darkgreen") +
  tm_shape(CalCOFI_bottle) +
  tm_dots(col = "blue") +
  tm_shape(lter_data_geom) + 
  tm_dots() +
  tm_shape(plumes_blooms) +
  tm_dots(col = "lightpink1") +
  tm_shape(USGS_2018) +
  tm_dots(col = "seashell") +
  tm_shape(USGS_2019) +
  tm_dots(col = "seashell") +
  tm_shape(LTER_bottle) +
  tm_dots(col = "purple")
```
