---
title: "Kelp Project Test"
author: "Jessica French"
date: "2023-01-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## First tutorial using `ncdf4` package following [this](https://rpubs.com/boyerag/297592) tutorial. 

```{r}
# load in packages
library(ncdf4)
library(tidync)
library(raster)
library(tmap)
```
## Current Question/ things I don't understand: 
**How to correctly set the extent, crs, and datum so the raster ends up in the correct spot. Right now it appears to cover the entire west coast with some of the highest kelp biomass in eastern Oregon...**

**Also don't understand how to access the time element/ layers. According to the metadata there should be measurement quarterly from 1984 to 2019 but I have only successfully gotten a single raster to display and all attempts to filter have not worked.**

first the output of the steps in the tutorial. 


```{r}
# # read in data used in tutorial
# tutorial_data <- nc_open(file.path(data_dir, 'gimms3g_ndvi_1982-2012.nc4'))
# 
# # print the metadata
# print(tutorial_data)
```


## Next the out put from the kelp file

link to data download: https://sbclter.msi.ucsb.edu/data/catalog/package/?package=knb-lter-sbc.74
```{r}
# Read in the netCDF file downloaded form the lter 
kelp_data <- nc_open('/Users/jfrench/Documents/MEDS/Capstone/DATA/GIMMS3G_NDVI_TRENDS_1275/data/LandsatKelpBiomass_2022_Q4_withmetadata.nc')

# print the metadata
print(kelp_data)
```
The rest of the outputs for the tutorial are shown on the webpage

```{r}
# extract the lat and long variables

lat <- ncvar_get(kelp_data, "latitude")
lon <- ncvar_get(kelp_data, "longitude")

area <- ncvar_get(kelp_data, "area")
```


```{r}
# From the metadata it looks like the location data is store in the station variable. Extract station variable and look at it 
 station <- ncvar_get(kelp_data, "station")

# When I have been able to look at the biomass variable as a data frame it appears to be all NAs so I am going to try to filter to any non NA values and if they exist. 
biomass <- ncvar_get(kelp_data, "biomass") 

# There are non-NA values
#biomass[is.na(biomass[,]) != TRUE]

# Try creating a slice in a time period I think will have values
biomass_slice <- biomass[,116]

# Try how they visualised in the tutorial
biomass_r <- raster(biomass, xmn=min(lon), xmx=max(lon), ymn=min(lat), ymx=max(lat), crs=CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs+ towgs84=0,0,0"))

area_r <- raster(area, xmn=min(lon), xmx=max(lon), ymn=min(lat), ymx=max(lat))
# Try plotting, it pops up but I don't really understand what I am looking at.
plot(area_r)

# lets try putting this on a tmap and see where is site relative to the channel islands
tmap_mode(mode = "view")
tm_shape(area_r) +
  tm_raster()


print(station)
print(biomass)
```
My interpretation of the situation is that the location information is is linked to the station variable. Each observation only has information for the station and each lat and lon has an associated station. Need to join the lat lon information to the observations based on station. 


```{r}
# lets look at the latitude variable
## extract using ncvar_get()
lat <- ncvar_get(kelp_data, "latitude")

# print
print(lat)
```

```{r}
# do the same thing with longitude

lon <- ncvar_get(kelp_data, "longitude")

print(lon)
```

